---
title: "Options, by package"
author: "Tim D. Smith"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Options, by package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(dplyr)
library(ggplot2)
library(magrittr)
library(reshape2)
library(viridis)

library(HomebrewGA)
invocations = fetch_install_invocations()
popularity = popularity_by_package(invocations)
opt_df = install_invocations_by_package(invocations)
```

# Top 100 packages

Total `install` invocations for top 100 packages, with and without formula options given; bar length indicates number of installs:

```{r popular_absolute, fig.height=10}
x = function() {
  sort_order = popularity %>% arrange(desc(n)) %>% head(100)
  popular = opt_df %>% filter(package %in% sort_order$package)
  popular %<>% inner_join(popularity) %>%
    filter(option == "none") %>%
    rename(without_option=count) %>%
    mutate(with_option=n-without_option) %>%
    melt(measure.vars=c("with_option", "without_option"),
         variable.name="option_given",
         value.name="count") %>%
    mutate(option_given=relevel(option_given, "without_option"),
           package = factor(package, levels=rev(sort_order$package))) %>%
    arrange(option_given)
  g = ggplot(popular, aes(package, count, fill=option_given)) +
    geom_bar(stat="identity") +
    coord_flip() +
    theme_bw()
  print(g)
}
x()
```

There are a few packages that are often installed with options; many of them offer `--universal` or multiple graphics backends.

# Top 100 packages, relative proportion

Relative proportion of `install` invocations with formula options given, for the top 100 packages; bar length indicates percentage of installs of each package:

```{r popular_relative, fig.height=10}
x = function() {
  sort_order = popularity %>% arrange(desc(n)) %>% head(100)
  popular = opt_df %>% filter(package %in% sort_order$package)
  popular %<>%
    inner_join(popularity) %>%
    filter(option == "none") %>%
    rename(without_option=count) %>%
    mutate(without_option=without_option/n, with_option=1-without_option) %>%
    melt(measure.vars=c("with_option", "without_option"),
         variable.name="option_given",
         value.name="count") %>%
    mutate(option_given=relevel(option_given, "without_option"),
           package = factor(package, levels=rev(sort_order$package))) %>%
    arrange(option_given)

  g = ggplot(popular, aes(package, count, fill=option_given)) +
    geom_bar(stat="identity") +
    scale_y_continuous(labels=scales::percent) +
    theme_bw() +
    coord_flip()
  print(g)
}
x()
```

# Fraction of installs with options vs. popularity

More popular packages are less likely to have many popular options. I would guess this is because more popular packages are less likely to offer options, but I haven't checked.

```{r n_downloads_vs_n_options_relative}
x = function() {
  options = opt_df %>%
    inner_join(popularity) %>%
    transform(fraction=count/n) %>%
    filter(option == "none") %>%
    mutate(fraction=1-fraction)
  g = ggplot(options, aes(n, fraction)) +
    geom_point(alpha=0.25) +
    geom_density2d(h=c(1,0.1)) +
    scale_x_log10() +
    labs(x="Total package installs", y="Fraction of installs per package with options") +
    theme_bw()
  print(g)
}
x()
```

The blue density contours show that nearly all packages are relatively unpopular and have very few installs with options, which is otherwise hard to see because of overplotting.

# Fraction of installs without options, by package

This shows that most packages are never installed with options. A handful of packages are very commonly installed with options.

```{r popularity_of_default_install_histogram}
x = function() {
  opt_df %<>%
    inner_join(popularity) %>%
    filter(option == "none") %>%
    mutate(default_fraction=count/n)
  g = ggplot(opt_df, aes(default_fraction)) +
    geom_histogram(binwidth=0.05) +
    scale_y_log10(breaks=c(1, 3, 10, 30, 100, 300, 1000, 3000, 10000)) +
    labs(x="Fraction of installs without options", y="Number of packages") +
    theme_bw()
  print(g)
}
x()
```

Similarly, packages often have a small number of absolute installations with 

```{r absolute_popularity_of_non_default_install_histogram}
x = function() {
  g = ggplot(opt_df %>% filter(option != "none"), aes(count, ..count..)) +
    geom_freqpoly() +
    scale_y_log10(breaks=c(1, 3, 10, 30, 100, 300, 1000, 3000, 10000)) +
    labs(x="Number of installs with options, per package", y="Number of packages") +
    theme_bw()
  print(g)
}
x()
```

# Packages with popular options (top 100)

The number on the left hand side indicates total popularity.

```{r packages_often_installed_with_options, fig.height=10}
x = function() {
  overall_popularity = popularity %>% rename(overall=n)
  merged = opt_df %>%
    filter(option == "none") %>%
    inner_join(overall_popularity) %>%
    filter(overall >= 100) %>%
    mutate(options_popularity=1-count/overall) %>%
    arrange(desc(options_popularity)) %>%
    head(100)
  
  merged$package %<>% factor(levels=rev(merged$package))
  g = ggplot(merged, aes(options_popularity, package)) +
    geom_point(aes(color=overall)) +
    xlim(0, 1) +
    geom_text(aes(size=overall, label=overall), x=0.1) +
    scale_color_viridis(trans="log10", guide="none") +
    scale_size_continuous(trans="log10", guide="none") +
    theme_bw()
  print(g)
}
x()
```
